### 2015-10-02 pics/blue.tex
add all figures

\begin{tikzpicture}[scale=1.0]

\begin{axis}[%scale=1.5,
             domain=0:12.0,
             restrict expr to domain={x}{0:12.0},
             %restrict expr to domain={y}{1.0E-12:7},
             xlabel={$E$ [eV]},
%             xmin=0,
%             xmax=1.5,
             ymin=0,
%             ymax=7,
             %xtick={30,50,...,170},
             %xticklabels={2,4,6,8,10,12,15,20,25},
             %ytick={-1.0,-0.8,...,1.0},
             %yticklabels={1.0,0.8,0.6,0.4,0.2,0.0,0.2,0.4,0.6,0.8,1.0},
             ylabel={$\Gamma$ [eV]},
             scale only axis,
             width=\textwidth-1.5cm,
             height=8cm,
             %ybar stacked
             axis line style = very thick,
             tick style = very thick
             ]

\addplot+[ycomb,
         mark=.,
         ultra thick,
         blue,
         forget plot
        ]
        table[
        x expr = \thisrowno{0},
        y expr = \thisrowno{1}
        ]
        {../data/blue_NeNe.dat};
        \addlegendimage{line legend, blue, very thick};
        \addlegendentry{set3 (blue)};

%\addplot[
%         mark=none,
%         diplom2,
%         ultra thick,
%         %dotted
%         ]
%         table[
%         x expr=\thisrowno{0},
%         y expr=\thisrowno{1}
%         ]
%        {../data/xear_3_surf.dat_intadd.inp.sp};
%         \addlegendentry{1 Xe, 55 Ar};
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{axis}
\end{tikzpicture}