### 2015-10-02 partners.tex
add all section files

\section{Decay with Decay Partners at Different Distances}