### 2015-10-05 computational.tex
add pictures and first introduction

\section{Computational Details}

\begin{table}[!h]
 \caption{Experimental values for the single ionization potentials
          \cite{Fasshauer14_1}
          used for the estimation of the decay widths.}
 \label{table:exp_input}
 \centering
 \begin{tabular}{lc}
  \toprule
  indicator            &  value \\
  \midrule
  SIP(Ne2s)            &  \unit[47.75]{eV} \\
  SIP(Ne2p)            &  \unit[21.10]{eV} \\
  SIP(Ar3p)$_{c<3}$    &  \unit[15.40]{eV} \\
  SIP(Ar3p)$_{c\ge 3}$ &  \unit[15.20]{eV} \\
  \bottomrule
 \end{tabular}
\end{table}