### 2015-10-05 theory.tex

\section{Theoretical Background}

\begin{align}
 E_{in}        &= SIP(X_{in}) \label{equation:E_in}\\          
 E_{fin}^\beta &= SIP(X_{D}^\beta) + SIP(X_{E}^\beta) + \frac 1d
           \label{equation:E_fin}\\                            
 E_{sec}^\beta &= E_{in}^\beta - E_{fin}^\beta \label{equation:E_sec}         
\end{align}

following Wentzel \cite{Wentzel27}, Feshbach\cite{Feshbach58,Feshbach62}
and Fano \cite{Fano61}

\begin{equation}
 \Gamma_{\beta}(E_{res}) = 2\pi \left|
                           \braket{\Phi_{in}| H_f |\chi_{\beta}}
                           \right|^2
\end{equation}