### 2015-11-30 pics/rot.tex

\begin{tikzpicture}[scale=1.0]

\begin{axis}[%scale=1.5,
             domain=0:12.0,
             restrict expr to domain={x}{0:12.0},
             %restrict expr to domain={y}{1.0E-12:7},
             xlabel={$E$ [eV]},
%             xmin=0,
%             xmax=1.5,
             ymin=0,
%             ymax=7,
             %xtick={30,50,...,170},
             %xticklabels={2,4,6,8,10,12,15,20,25},
             %ytick={-1.0,-0.8,...,1.0},
             %yticklabels={1.0,0.8,0.6,0.4,0.2,0.0,0.2,0.4,0.6,0.8,1.0},
             ylabel={$\Gamma$ [eV]},
             scale only axis,
             width=\textwidth-1.5cm,
             height=8cm,
             %ybar stacked
             axis line style = very thick,
             tick style = very thick
             ]

\addplot+[ycomb,
         mark=.,
         ultra thick,
         red,
         forget plot
        ]
        table[
        x expr = \thisrowno{0},
        y expr = \thisrowno{1}
        ]
        {../data/red_NeNe.dat};
        \addlegendimage{line legend, red, very thick};
        \addlegendentry{set 5 (red)};

\draw[thick] (axis cs:1.40,0.007) arc [radius=0.77cm,start angle=10,end angle=170];
\node[pin={[pin distance=0.2cm]90:{\small \textbf{(a)}\\ across shells}}]
     at (axis cs:1.0,0.0080) {};

\draw[thick] (axis cs:1.90,0.0120) arc [radius=0.55cm,start angle=10,end angle=170];
\node[pin={[pin distance=0.2cm]10:{\small \textbf{(b)} same shell}}]
     at (axis cs:1.9,0.0125) {};

\draw[thick] (axis cs:2.90,0.0005) arc [radius=0.50cm,start angle=10,end angle=170];
\node[pin={[pin distance=0.2cm]92:{\small \textbf{(c)} across shells}}]
     at (axis cs:2.70,0.0010) {};

\end{axis}
\end{tikzpicture}