### 2016-01-10 letters/reconsider.tex

\documentclass[DIN,10pt,pagenumber=false,parskip=half,fromalign=left,fromphone=false,fromemail=true,fromurl=false,fromlogo=true,fromrule=false]{scrlttr2}
\usepackage[utf8]{inputenc}
\usepackage{ngerman}
\usepackage{units}
\usepackage{tabularx}
\usepackage[right]{eurosym}
\usepackage{graphicx}
\usepackage[top=2cm,bottom=0cm,left=2cm,right=2cm]{geometry}
%\RequirePackage{graphicx}

\setkomavar{fromname}{Dr. Elke Faßhauer}
\setkomavar{fromaddress}{CTCC\\ University of Tromsø{}
                         --- The Arctic University of Norway
                         \\ 9037 Tromsø \\ Norway}
%\setkomavar{fromphone}{06221/545220}
\setkomavar{fromemail}{elke.fasshauer@uit.no}
\setkomavar{subject}{Manuscript AK11318}
\setkomavar{signature}{Elke Faßhauer -- author of NJP-104327}
\setkomavar{fromlogo}{\includegraphics[width=3cm]{uit.pdf}}

\begin{document}

\begin{letter}{New Journal of Physics\\Editorial Board}
\textheight200mm
\opening{Dear Editor,}

I would kindly like you to reconsider your evaluation of my manuscript
NJP-104327 submitted to New Journal of Physics.

There where two critizied aspects of my manuscript: 1. it being mostly directed
towards the ICD community and 2. the topic to be subtle and not conceptually new.
First, the manuscript is directed towards the ICD community but furthermore also
to the cluster and condensed matter community and indirectly to the cancer
research community. It generalizes earlier descriptions of non-nearest neighbour
ICD which is not a subtle topic in itself which we shortly mentioned earlier
(New Journal of Physics \textbf{16}, 103026 (2014).), since it explains peaks in the spectrum the
large shoulders in the ICD spectra of clusters and intrinsically carries
information about the cluster structure.

One of the big current topics is to describe the DNA damage following an ICD in
water molecules around it. Obviously, in this setting several water molecules
at different distances are involved. Understanding the effect of non-nearest
neighbour ICD is crucial for the theoretical description of the process and the
interpretation of experimental spectra. My manuscript provides the basis for
this research.

I am convinced that my article is a valuable contribution for the readership of New Journal of Physics, highly appreciated by scientists considering Interatomic Coulombic
Decay (ICD), causations of cancer as well as
the transition from atomic and molecular clusters to the solid state and beyond.

I kindly ask you to forward my work to peer review.\\

\closing{Yours sincerely,}
\end{letter}

\end{document}