### 2016-01-10 theory.tex

\section{Theoretical Background}
\label{sec:theory}

In order to simulate the ICD electron spectra for a given cluster structure,
the kinetic energies of the ICD electrons $E_{ICD}$ and the corresponding
decay widths of all pairs have to be determined. The ICD electron energies
are given by the differences between the initial state and the final state
energies $E_{in}$ and $E_{fin}$, respectively. The initial state energy
is given by the single ionization potential (SIP) of the sub-valence electron
of the entire system
and the final state energy is given by the double ionization potential (DIP).
In the asymptotic limit, which is a reasonably good approximation for
weakly bound systems, the initial state energy is approximated by the SIP
of the initially ionized unit $X_{in}$ and the final state energy can be
approximated
by the sum over the SIPs of the electron donating unit $X_D$ and the electron
emitting unit $X_E$ ionized in the final state as well as the
Coulomb repulsion between two point charges at the interatomic distance $R$

\begin{eqnarray}
 E_{ICD}^\beta &= E_{in}^\beta - E_{fin}^\beta \label{equation:E_sec}\\
 E_{in}        &= SIP(X_{in}) \label{equation:E_in}\\          
 E_{fin}^\beta &= SIP(X_{D}^\beta) + SIP(X_{E}^\beta) + \frac 1R
           \label{equation:E_fin}                . 
\end{eqnarray}
Here, $\beta$ denotes the selected decay channel.

Following Wentzel \cite{Wentzel27}, Feshbach\cite{Feshbach58,Feshbach62}
and Fano \cite{Fano61} the decay width is given by

\begin{equation}
 \Gamma_{\beta}(E_{res}) = 2\pi \left|
                           \braket{\Phi_{in}| H_f |\chi_{\beta}}
                           \right|^2   .
\end{equation}
Here, the bound and ionized initial state is described by $\ket{\Phi_{in}}$ and
the final continuum state of a particular decay channel $\beta$ is
given by $\ket{\chi_{\beta}}$.
Its challenging description involving both bound and continuum states can
amongst others be achieved by the FanoADC-Stieltjes approach, where a
subset of the  $2h1p$ (2-hole-1-particle) functions within the
Algebraic Diagrammatic Construction
(ADC) are used to mimic the final state function $\ket{\chi_{\beta}}$.
By these means calculated discrete energies and corresponding transitions
moments are then used to construct a continuous function, which is evaluated
at the resonance energy approximated by the single ionization potential
corresponding to the initial state $E_{in}$. For a more detailed
description of the method and comparison to other approaches see References
\cite{Averbukh05,Fasshauer15_1} and references therein.