### 2016-03-17 revised/NJP_master.tex
format of the paper - prepare submission

%\documentclass[aps,amssymb,preprint,a4paper,dvips]{revtex4}
\documentclass[12pt,twocolumn]{iopart}

%\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{braket}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{placeins}
%\usepackage{amsmath,amsfonts,amssymb}
\usepackage{iopams}
\usepackage{dsfont}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{units}
%\usepackage{natbib}

%Finde die Bilder
%\graphicspath{{./pics/}}

% Alphabetic counting in footnotes
\renewcommand{\thefootnote}{\alph{footnote}}

% Captions over full width

%\usepackage{caption}
%\captionsetup[figure]{format=plain, justification=raggedright, singlelinecheck=false, font=footnotesize, labelfont=bf}
%\captionsetup[table]{position=top, format=plain, justification=raggedright, singlelinecheck=false, font=footnotesize, labelfont=bf}

\begin{document}

\setlength{\tabcolsep}{12pt}

% Define some colour
\definecolor{diplom1}{rgb}{0.0 0.4 1.0}
\definecolor{diplom2}{rgb}{0.0 0.0 0.6}
\definecolor{diplom3}{RGB}{153,0,0} %unirot

%\title[Non-nearest Neighbour ICD in Clusters]{Non-nearest Neighbour ICD in Clusters: A Way to Distinguish Icosahedral
%       and Cuboctahedral Clusters?}
\title[Non-nearest Neighbour ICD in Clusters]{Non-nearest Neighbour ICD in
       Clusters}

\author{E Fasshauer}
\address{Centre for Theoretical and Computational Chemistry,
         Department of Chemistry, University of Troms\o,
         -- The Arctic University of Norway, N-9037 Troms\o, Norway}

\eads{\mailto{elke.fasshauer@uit.no}}

\date{\today}

\begin{abstract}
 \input{abstract}
\end{abstract}

\pacs{33.50.Hv, 33.70.Fd, 33.70.Jg, 34.80.Dp, 36.40.-c, 36.40.Mr}
\submitto{New Journal of Physics}

\maketitle

\input{intro}
\input{theory}
\input{computational}
\input{partners}
\input{clusters}
\input{summary}

\input{ack}

\eject
\section*{References}
%\bibliographystyle{unsrt}
%\bibliography{theolit}

\begin{thebibliography}{10}

\bibitem{Hergenhahn11}
U.~Hergenhahn.
\newblock {Interatomic and Intermolecular Coulombic Decay: The Early Years}.
\newblock {\em J. Electron Spectrosc. Relat. Phenom.}, 184:78, 2011.

\bibitem{Jahnke15}
T.~Jahnke.
\newblock {Interatomic and intermolecular Coulombic decay: the coming of age
  story}.
\newblock {\em J. Phys. B: Atomic, Molecular and Optical Physics}, 48:082001,
  2015.

\bibitem{Cederbaum97}
L.~S. Cederbaum, J.~Zobeley, and F.~Tarantelli.
\newblock {Giant Intermolecular Decay and Fragmentation of Clusters}.
\newblock {\em Phys. Rev. Lett.}, 79(24):4778, 1997.

\bibitem{Marburger03}
S.~Marburger, O.~Kugeler, U.~Hergenhahn, and T.~M{\"o}ller.
\newblock Experimental evidence for interatomic coulombic decay in ne clusters.
\newblock {\em Phys. Rev. Lett.}, 90(20):203401, 2003.

\bibitem{Ohrwall04}
G.~{\"O}hrwall, M.~Tchaplyguine, M.~Lundwall, R.~Feifel, H.~Bergersen,
  T.~Rander, A.~Lindblad, J.~Schulz, S.~Peredkov, S.~Barth, S.~Marburger,
  U.~Hergenhahn, S.~Svensson, and O.~Bj{\"o}rneholm.
\newblock {Femtosecond Interatomic Coulombic Decay in Free Neon Clusters: Large
  Lifetime Differences between Surface and Bulk}.
\newblock {\em Phys. Rev. Lett.}, 93(17):173401, 2004.

\bibitem{Jahnke04}
T.~Jahnke, A.~Czasch, M.~S. Sch{\"o}ffler, S.~Sch{\"o}ssler, A.~Knapp,
  M.~K{\"a}sz, J.~Titze, C.~Wimmer, K.~Kreidi, R.~E. Grisenti, A.~Staudte,
  O.~Jagutzki, U.~Hergenhahn, H.~Schmidt-B{\"o}cking, and R.~D{\"o}rner.
\newblock Experimental observation of interatomic coulombic decay in neon
  dimers.
\newblock {\em Phys. Rev. Lett.}, 93(16):163401, 2004.

\bibitem{Santra00_1}
R.~Santra, J.~Zobeley, L.~S. Cederbaum, and N.~Moiseyev.
\newblock {Interatomic Coulombic Decay in van der Waals Clusters and Impact of
  Nuclear Motion}.
\newblock {\em Phys. Rev. Lett.}, 85:4490, 2000.

\bibitem{Santra01_1}
R.~Santra, J.~Zobeley, L.~S. Cederbaum, and F.~Tarantelli.
\newblock {Intermolecular Coulombic decay of clusters}.
\newblock {\em J. Electron Spectrosc. Relat. Phenom.}, 114--116:41, 2001.

\bibitem{Santra01_3}
R.~Santra, J.~Zobeley, and L.~S. Cederbaum.
\newblock {Electronic decay of valence holes in clusters and condensed matter}.
\newblock {\em Phys. Rev. B}, 64(24):245104, 2001.

\bibitem{Averbukh06_1}
V.~Averbukh and L.~S. Cederbaum.
\newblock {Calculation of interatomic decay widths of vacancy states
  delocalized due to inversion symmetry}.
\newblock {\em J. Chem. Phys.}, 125:094107, 2006.

\bibitem{Lundwall07}
M.~Lundwall, W.~Pokapanich, H.~Bergersen, A.~Lindblad, T.~Rander, G.~\"Ohrwall,
  M.~Tchaplyguine, S.~Barth, U.~Hergenhahn, S.~Svensson, and O.~Bj\"orneholm.
\newblock Self-assembled heterogeneous argon/neon core-shell clusters studied
  by photoelectron spectroscopy.
\newblock {\em J. Chem. Phys.}, 126:214706, 2007.

\bibitem{Jahnke10}
T.~Jahnke, H.~Sann, T.~Havermeier, K.~Kreidi, C.~Stuck, M.~Meckel,
  M.~Sch{\"o}ffler, N.~Neumann, R.~Wallauer, S.~Voss, A.~Czasch, O.~Jagutzki,
  A.~Malakzadeh, F.~Afaneh, Th. Weber, H.~Schmidt-B{\"o}cking, and
  R.~D{\"o}rner.
\newblock {Ultrafast energy transfer between water molecules}.
\newblock {\em Nature Physics}, 6:139, 2010.

\bibitem{Forstel13}
M.~F{\"o}rstel, T.~Arion, and U.~Hergenhahn.
\newblock {Measuring the efficiency of interatomic Coulombic decay in Ne
  clusters}.
\newblock {\em J. Electron Spectrosc. Relat. Phenom.}, 191:16, 2013.

\bibitem{Mucke15}
M.~Mucke, T.~Arion, M.~F{\"o}rstel, T.~Lischke, and U.~Hergenhahn.
\newblock {Competition of inelastic electron scattering and Interatomic
  Coulombic Decay in Ne clusters}.
\newblock {\em J. Electron Spectrosc. Relat. Phenom.}, 200:232, 2015.

\bibitem{Fasshauer14_1}
E.~Fasshauer, M.~F\"orstel, S.~Pallmann, M.~Pernpointner, and U.~Hergenhahn.
\newblock Using icd for structure analysis of clusters: a case study on
  coexpanded ne-ar clusters.
\newblock {\em New J. Phys.}, 16:103026, 2014.

\bibitem{Zobeley98}
J.~Zobeley, L.~S. Cederbaum, and F.~Tarantelli.
\newblock Highly excited electronic states of molecular clusters and their
  decay.
\newblock {\em J. Chem. Phys.}, 108(23):9737, 1998.

\bibitem{Pernpointner06}
M.~Pernpointner, S.~Knecht, and L.~S. Cederbaum.
\newblock {Ionization spectra and electronic decay in small iodide clusters:
  Fully relativistic results}.
\newblock {\em J. Chem. Phys.}, 125:034309, 2006.

\bibitem{Mueller06}
I.~B. M{\"u}ller and L.~S. Cederbaum.
\newblock {Ionization and double ionization of small water clusters}.
\newblock {\em J. Chem. Phys.}, 125:204305, 2006.

\bibitem{Kryzhevoi11_1}
N.~V. Kryzhevoi and L.~S. Cederbaum.
\newblock {Using pH Value To Control Intermolecular Electronic Decay}.
\newblock {\em Angew. Chem. Int. Ed.}, 50:1306, 2011.

\bibitem{Ghosh13a}
A.~Ghosh, S.~Pal, and N.~Vaval.
\newblock {Study of interatomic Coulombic decay of Ne(H2O)n (n=1,3) clusters
  using equation-of-motion coupled-cluster method}.
\newblock {\em J. Chem. Phys.}, 139:064112, 2013.

\bibitem{Bande13}
A.~Bande.
\newblock Electron dynamics of interatomic coulombic decay in quantum dots
  induced by a laser field.
\newblock {\em J. Chem. Phys.}, 138:214104, 2013.

\bibitem{Harbach13}
P.~H.~P. Harbach, M.~Schneider, S.~Faraji, and A.~Dreuw.
\newblock Intermolecular coulombic decay in biology: the initial electron
  detachment from fadh- in dna photolyases.
\newblock {\em J. Phys. Chem. Lett.}, 4:943, 2013.

\bibitem{Alizadeh15}
E.~Alizadeh, T.~M. Orlando, and L.~Sanche.
\newblock {Biomolecular Damage Induced by Ionizing Radiation: The Direct and
  Indirect Effects of Low-Energy Electrons on DNA}.
\newblock {\em Annu. Rev. Phys. Chem.}, 66:379, 2015.

\bibitem{Gokhberg14}
K.~Gokhberg, P.~Koloren\v{c}, A.~I. Kuleff, and L.~S. Cederbaum.
\newblock {Site- and energy-selective slow-electron production through
  intermolecular Coulombic decay}.
\newblock {\em Nature}, 505:661, 2014.

\bibitem{Trinter14}
F.~Trinter, M.~S. Sch{\"o}ffler, H.~K. Kim, F.~P. Sturm, K.~Cole, N.~Neumann,
  A.~Vredenborg, J.~Williams, I.~Bocharova, R.~Guillemin, M.~Simon,
  A.~Belkacem, A.~L. Landers, Th. Weber, H.~Schmidt-B{\"o}cking, R.~D{\"o}rner,
  and Jahnke T.
\newblock {Resonant Auger decay driving intermolecular Coulombic decay in
  molecular dimers}.
\newblock {\em Nature}, 505:664, 2014.

\bibitem{Fasshauer13}
E.~Fasshauer, M.~Pernpointner, and K.~Gokhberg.
\newblock Interatomic decay of inner-valence ionized states in arxe clusters:
  Relativistic approach.
\newblock {\em J. Chem. Phys.}, 138:014305, 2013.

\bibitem{kryzhevoi2007}
N.~V. Kryzhevoi, V.~Averbukh, and L.~S. Cederbaum.
\newblock High activity of helium droplets following ionisation of systems
  inside those droplets.
\newblock {\em Phys. Rev. B}, 76:094513, 2007.

\bibitem{Fasshauer16_1}
E.~Fasshauer, M.~F\"orstel, M.~M\"ucke, M.~Pernpointner, and U.~Hergenhahn.
\newblock in preparation.

\bibitem{Hagena72}
O.~F. Hagena and W.~Obert.
\newblock Cluster formation in expanding supersonic jets: Effect of pressure,
  temperature, nozzle size, and test gas.
\newblock {\em The Journal of Chemical Physics}, 56(5):1793--1802, 1972.

\bibitem{Hagena81}
O.~F. Hagena.
\newblock Nucleation and growth of clusters in expanding nozzle flows.
\newblock {\em Surf. Sci.}, 106:101, 1981.

\bibitem{Hagena87}
O.~F. Hagena.
\newblock Condensation in free jets: Comparison of rare gases and metals.
\newblock {\em Z. Phys. D}, 4:291, 1987.

\bibitem{Benfield92}
R.~E. Benfield.
\newblock Mean coordination numbers and the non-metal-metal transition in
  clusters.
\newblock {\em J. Chem. Soc. Faraday Trans.}, 88:1107, 1992.

\bibitem{Hartke02}
B.~Hartke.
\newblock Structural transitions in clusters.
\newblock {\em Angew. Chem. Int. Ed.}, 41:1468, 2002.

\bibitem{Pahl08}
E.~Pahl, F.~Calvo, L.~Ko\v{c}i, and P.~Schwerdtfeger.
\newblock Accurate melting temperatures for neon and argon from ab initio monte
  carlo simulations.
\newblock {\em Angew. Chem. Int. Ed.}, 47:8207, 2008.

\bibitem{Wentzel27}
G.~Wentzel.
\newblock {\"U}ber strahlungslose quantenspr{\"u}nge.
\newblock {\em Z. Physik}, 43:524, 1927.

\bibitem{Feshbach58}
H.~Feshbach.
\newblock Unified theory of nuclear reactions.
\newblock {\em Ann. Phys.}, 5:357, 1958.

\bibitem{Feshbach62}
H.~Feshbach.
\newblock A unified theory of nuclear reactions. ii.
\newblock {\em Ann. Phys.}, 19:287, 1962.

\bibitem{Fano61}
U.~Fano.
\newblock Effects of configuration interaction on intensities and phase shifts.
\newblock {\em Phys. Rev.}, 124(6):1866--1878, Dec 1961.

\bibitem{Averbukh05}
V.~Averbukh and L.~S. Cederbaum.
\newblock Ab initio calculation of interatomic decay rates by a combination of
  the fano ansatz, green's-function methods, and the stieltjes imaging
  technique.
\newblock {\em J. Chem. Phys.}, 123:204107, 2005.

\bibitem{Fasshauer15_1}
E.~Fasshauer, P.~Koloren\ifmmode~\check{c}\else \v{c}\fi{}, and
  M.~Pernpointner.
\newblock Relativistic decay widths of autoionization processes: The
  relativistic fanoadc-stieltjes method.
\newblock {\em J. Chem. Phys.}, 142:144106, 2015.

\bibitem{Bondi64}
A.~Bondi.
\newblock van der waals volumes and radii.
\newblock {\em J. Phys. Chem.}, 68:441, 1964.

\bibitem{HARDRoC}
{HARDRoC}, Hunting Asymptotic Relativistic Decay Rates of Clusters (2013),
  written by E.~Fasshauer (see
  http://www.pci.uni-heidelberg.de/tc/usr/elke/hardroc/html/main.html).

\bibitem{Fasshauer_thesis}
E.~Fasshauer.
\newblock {\em Investigation of Relativistic Effects in Electronic Decay
  Processes in Small and Large Noble Gas Clusters by Ab Initio and New
  Simulation Approaches}.
\newblock PhD thesis, University of Heidelberg, 2014.

\bibitem{Scheit06}
S.~Scheit, V.~Averbukh, H.-D. Meyer, J.~Zobeley, and L.~S. Cederbaum.
\newblock {Interatomic Coulombic decay in a heteroatomic rare gas cluster}.
\newblock {\em J. Chem. Phys.}, 124:154305, 2006.

\bibitem{OKeeffe14}
P.~O'Keeffe, A.~Ciavardini, E.~Ripani, P.~Bolognesi, M.~Coreno, L.~Avaldi,
  M.~Devetta, M.~Di~Fraia, C.~Callegari, K.~C. Prince, and R.~Richter.
\newblock {Experimental investigation of the interatomic Coulombic decay in
  NeAr dimers}.
\newblock {\em Phys. Rev. A}, 90:042508, 2014.

\bibitem{Schnorr13}
K.~Schnorr, A.~Senftleben, M.~Kurka, A.~Rudenko, L.~Foucar, G.~Schmid,
  A.~Broska, T.~Pfeifer, K.~Meyer, D.~Anielski, R.~Boll, D.~Rolles,
  M.~K{\"u}bel, M.~F. Kling, Y.~H. Jiang, S.~Mondal, T.~Tachibana, K.~Ueda,
  T.~Marchenko, M.~Simon, G.~Brenner, R.~Treusch, S.~Scheit, V.~Averbukh,
  J.~Ullrich, C.~D. Schr{\"o}ter, and R.~Moshammer.
\newblock {Time-Resolved Measurement of Interatomic Coulombic Decay in Ne$_2$}.
\newblock {\em Phys. Rev. Lett.}, 111:093402, 2013.

\bibitem{Schnorr15}
K.~Schnorr, A.~Senftleben, G.~Schmid, S.~Augustin, M.~Kurka, A.~Rudenko,
  L.~Foucar, A.~Broska, K.~Meyer, D.~Anielski, D.~Anielski, R.~Boll, D.~Rolles,
  M.~K{\"u}bel, M.~F. Kling, Y.~H. Jiang, S.~Mondal, T.~Tachibana, K.~Ueda,
  T.~Marchenko, M.~Simon, G.~Brenner, R.~Treusch, S.~Scheit, V.~Averbukh,
  J.~Ullrich, T.~Pfeifer, C.~D. Schr{\"o}ter, and R.~Moshammer.
\newblock {Time-resolved study of ICD in Ne dimers using FEL radiation}.
\newblock {\em J. Electron Spectrosc. Relat. Phenom.}, 204:245, 2015.

\bibitem{Barth06_2}
S.~Barth, S.~Marburger, O.~Kugeler, V.~Ulrich, S.~Joshi, A.~M. Bradshaw, and
  U.~Hergenhahn.
\newblock {The efficiency of Interatomic Coulombic Decay in Ne clusters}.
\newblock {\em Chem. Phys.}, 329:246, 2006.

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}