### 2016-03-17 revised/abstract.tex

\begin{abstract}
Interatomic Coulombic Decay (ICD) is an electronic decay process of
excited, ionized systems. It has been shown to occur in a multitude of small
and large systems.
The effects of more than one possible decay
partner are discussed in detail illustrated by simulated ICD electron spectra
of NeAr clusters and pure Ne clusters.
Hereby, the mostly underestimated contribution of decay with
non-nearest neighbours is highlighted. In the neon clusters, the lifetime of the
bulk atoms is found to be in excellent agreement with experiment
(\emph{Phys. Rev. Lett.}, \textbf{93}, 173401 (2004).), while the
lifetimes of the surface atoms differ significantly. Hence, the experimental
lifetime can not purely be explained by the effect of the number of
neighbours.

We propose the possibility to investigate the transition from
small clusters to the solid state by using the ICD electron spectra to
distinguish between icosahedral and cuboctahedral cluster structures.
\end{abstract}