### 2016-03-17 revised/computational.tex

\section{Computational Details}
\label{sec:computational}
The cluster structures used were constructed to have an ideal icosahedral or
face-centered-cubic geometry with optional additional incomplete outermost
shells. These are based on the van der Waals radii for neon
$r_{Ne}=$ \unit[1.54]{\AA}
and argon $r_{Ar}=$ \unit[1.88]{\AA} \cite{Bondi64}.
In case of the NeAr clusters two of those
cluster structures from Ref. \cite{Fasshauer14_1}, where the theoretical
and experimental argon content in the cluster,
the NeAr-ICD to total
ICD ratio and the peak position of the NeAr-ICD peak matched best for a given
manifold (set) of clusters produced under certain experimental conditions,
have been chosen.
These are $C_{Ar}=3$, $C_{Ne}=1$, $S_{Ne}=7$ for set 3
and $C_{Ar}=2$, $C_{Ne}=1$, $S_{Ne}=13$ for set 5 where we used the same
nomenclature as in Ref. \cite{Fasshauer14_1}. Hence, $C_{Ar}$ denotes
the edge length of the argon core, $C_{Ne}$ denotes the number of complete
neon shells around the argon core and $S_{Ne}$ denotes the number of 
triangular surfaces additionally covered by neon atoms.

\begin{table}[h]
 \caption{Experimental values for the single ionization potentials
          \cite{Fasshauer14_1}
          used for the estimation of the decay widths.}
 \label{table:exp_input}
 \centering
 \begin{tabular}{lc}
  \toprule
  indicator            &  value \\
  \midrule
  SIP(Ne2s)            &  \unit[47.75]{eV} \\
  SIP(Ne2p)            &  \unit[21.10]{eV} \\
  SIP(Ar3p)$_{c<3}$    &  \unit[15.40]{eV} \\
  SIP(Ar3p)$_{c\ge 3}$ &  \unit[15.20]{eV} \\
  \bottomrule
 \end{tabular}
\end{table}

The calculations to obtain the ICD electron spectra of all cluster
structures were performed with
the program HARDRoC \cite{HARDRoC} based on the decomposition of the cluster into
pairs. Every pair with the same distance is treated equally. This includes the
assumption that every neon atom is ionized with the same probability.
\cite{Fasshauer14_1,Fasshauer_thesis}
It uses experimental ionization energies shown
in Table \ref{table:exp_input} and curves fitted to the decay width of the
NeNe ICD of Ref. \cite{Averbukh06_1} and the NeAr decay widths of Ref.
\cite{Fasshauer14_1}. These fitted curves yield lifetimes
$\tau=\frac{\hbar}{\Gamma}$ at the corresponding equilibrium distances of
$\tau_{NeNe}= \unit[60]{fs}$ and
$\tau_{NeAr}= \unit[44]{fs}$.